Scheduling and Amortization for MPC

### 2022-01-30 paper_SIMD.tex
Sec. 4.5 Work in progress.

\ana{Commented out previous writeup. Removal/handling of targetless phi-nodes will go into the Extension to Basic vectorization.}

@@

\ana{TODO: Handling of Targetless phi nodes should be done here, as an extension to Basic Vectorization.}

@@

The other change concerns def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable (e.g., the definition can be a propagation $A_2 = A_1$ or a pseudo $\phi$-node $A_2 = \phi(A_0,A_1)$). These edges are not handled in the same way as in Basic Vectorization, specifically, we do not expand dimension as we do for scalars in Basic Vectorization according to the loop nest. A key invariant is that the dimension of an array $A$ cannot drop below $A$'s canonical dimensionality. Below we enumerate the cases of def-use edges.

@@

Recall that after removal of infeasible edges and redundant phi-nodes, the Aiken's array write example will be (roughly) as follows:

\begin{algorithmic}
\STATE for i in range(N):
\STATE 1. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
\STATE 2. $\texttt{A}_1$ = $\texttt{A}$; $\texttt{A}_1$[i] = $\texttt{B}$[i] + 10;
\STATE 3. $\texttt{B}_1$ = $\texttt{B}$; $\texttt{B}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
\STATE 4. $\texttt{C}_1$ = $\texttt{C}$; $\texttt{C}_1$[i] = $\texttt{A}_1$[i] * $\texttt{D}_0$[i-1];
\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1$[i] = $\texttt{B}_1$[i] * $\texttt{C}_1$[i];
\end{algorithmic}

There are no nested loops, thus, array accesses remain as is.

After Phase 1 of Basic vectorization we have the following code:
\begin{algorithmic}
\STATE for i in range(N):
\STATE 1. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
\STATE 2. $\texttt{A}_1$ = $\texttt{A}$; $\texttt{A}_1[\stackrel{\rightarrow}{i}] = \texttt{B}[\stackrel{\rightarrow}i] + [10,...]$
\STATE 3. $\texttt{B}_1$ = $\texttt{B}$; $\texttt{B}_1[\stackrel{\rightarrow}i] = \texttt{A}_1[\stackrel{\rightarrow}i] * \texttt{D}_0[\stackrel{\rightarrow}{i-1}]$
\STATE 4. $\texttt{C}_1$ = $\texttt{C}$; $\texttt{C}_1[\stackrel{\rightarrow}i] = \texttt{A}_1[\stackrel{\rightarrow}i] * \texttt{D}_0[\stackrel{\rightarrow}{i-1}]$
\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1[\stackrel{\rightarrow}i] = \texttt{B}_1[\stackrel{\rightarrow}i] * \texttt{C}_1[\stackrel{\rightarrow}i]$
\end{algorithmic}

>>>>

%\ana{Commented out previous writeup. Removal/handling of targetless phi-nodes will go into the Extension to Basic vectorization.}

@@

\ana{TODO: This section is work in progress!}

@@

The other change concerns def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable (e.g., the definition can be an update $A_2 = update(A_1,...)$ or a pseudo $\phi$-node $A_2 = \phi(A_0,A_1)$). These edges are not handled in the same way as in Basic Vectorization, specifically, we do not expand dimension as we do for scalars in Basic Vectorization according to the loop nest. A key invariant is that the dimension of an array $A$ cannot drop below $A$'s canonical dimensionality. Below we enumerate the cases of def-use edges.

@@

Recall that after removal of infeasible edges, the Aiken's array write example will be (roughly) as follows:

\begin{algorithmic}
\STATE for i in range(N):
\STATE 1. $\texttt{A}_1$ = $\phi(\texttt{A}_0,\texttt{A}_2)$
\STATE 2. $\texttt{B}_1$ = $\phi(\texttt{B}_0,\texttt{B}_2)$
\STATE 3. $\texttt{C}_1$ = $\phi(\texttt{C}_0,\texttt{C}_2)$
\STATE 4. $\texttt{D}_2$ = $\phi(\texttt{D}_0,\texttt{D}_2)$
\STATE 5. $\texttt{A}_2$ = update($\texttt{A}_1$, i, $\texttt{B}_1$[i] + 10);
\STATE 6. $\texttt{B}_2$ = update($\texttt{B}_1$, i, $\texttt{A}_2$[i] * $\texttt{D}_1$[i-1]);
\STATE 7. $\texttt{C}_2$ = update($\texttt{C}_1$, i, $\texttt{A}_2$[i] * $\texttt{D}_1$[i-1]);
\STATE 8. $\texttt{D}_2$ = update($\texttt{D}_1$; i, $\texttt{B}_2$[i] * $\texttt{C}_2$[i]);
\end{algorithmic}

There are no nested loops, thus, array accesses remain as is. The closure of 4 includes 4,6,7, and 8, 
while 5 is vectorizable.

Removal of infeasible edges leaves $\phi$-nodes 1,2, and 3, i.e., A,B, and C as targetless. 
After Phase 1 of Basic vectorization we have the following code:
\begin{algorithmic}
\STATE for i in range(N):
\STATE 1. $\texttt{D}_0$ = $\phi(\texttt{D},\texttt{D}_1)$
\STATE 2. $\texttt{A}_1$ = $\texttt{A}$; $\texttt{A}_1[\stackrel{\rightarrow}{i}] = \texttt{B}[\stackrel{\rightarrow}i] + [10,...]$
\STATE 3. $\texttt{B}_1$ = $\texttt{B}$; $\texttt{B}_1[\stackrel{\rightarrow}i] = \texttt{A}_1[\stackrel{\rightarrow}i] * \texttt{D}_0[\stackrel{\rightarrow}{i-1}]$
\STATE 4. $\texttt{C}_1$ = $\texttt{C}$; $\texttt{C}_1[\stackrel{\rightarrow}i] = \texttt{A}_1[\stackrel{\rightarrow}i] * \texttt{D}_0[\stackrel{\rightarrow}{i-1}]$
\STATE 5. $\texttt{D}_1$ = $\texttt{D}_0$; $\texttt{D}_1[\stackrel{\rightarrow}i] = \texttt{B}_1[\stackrel{\rightarrow}i] * \texttt{C}_1[\stackrel{\rightarrow}i]$
\end{algorithmic}
